Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-03-10 index.qmd
3/10 updats

@@

the `cox.zph()` calculates the scaled Schoenfeld partial residuals and
returns the $\chi^2$ value, the degrees of freedom, and p-values for
each individual covariate along with a global value. Using an
$\alpha = 0.05$ and the global p-value, the determination of whether to
reject or fail to reject $H_0$ can be made [@nahhas2025].

@@

@@

@@

A forest plot will be created using the `plot_model()` function from the sjPlot package. The CPH model can be passed as the sole argument to visualize the hazard ratios.

@@

After generating the CPH model, a Kaplan-Meier survival curve can be generated to view survival probabilities against time. These plots are important when modeling survival data because they clearly demonstrate the time where the event being modeled for is expected to occur.

The survival curve will be plotted by using the survival package's `survfit()` function. The object created will be plotted using `plot()`

@@

As defined by the formula, cumulative hazard is a summation of the hazard rates for an individual from time 0 to time $t$. A higher cumulative hazard indicates a higher risk of event. Cumulative hazard is used as an additional measure for risk, but it is mainly used to calculate survival probability to plot against time on survival curves.

The cumulative hazard will be calculated in R by grouping the data by the id field and finding the summation of the hazard rates for the individual through the duration of the observation period. A column will be added showing the cumulative hazard for each individual at ach point in time.

@@

The resulting survival probability will be between 0 and 1 with a value close to one indicating a high probability of the individual not experiencing the event. Conversely, a value close to 0 indicates a higher probability of the individual experiencing the event. A survival probability of .05 indicates that only 5% of individuals under the same conditions would survive. Survival probability is an important metric when modeling, visualizing, and making predictions with survival data.

@@

```{r, echo = FALSE, warning=FALSE,include=FALSE}
#Load packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)
library(knitr)
library(sjPlot)

@@

```{r}
survfit <- survfit(cox, newdata = test)
plot(survfit, 
     col = "blue", 
     conf.int = TRUE, 
     mark.time = FALSE,
     xlab = "Time (Iterations)",
     ylab = "Survival Probability",
     main = "Survival Probability Over Time")
```

>>>>

```{r, echo = FALSE, warning=FALSE, include=FALSE}
#Load packages
library(tidyverse)
library(gridExtra)
library(survival)
library(survminer)
library(ggpubr)
library(magrittr)
library(ggfortify)
library(knitr)
library(sjPlot)

@@

the `cox.zph()` calculates the scaled Schoenfeld partial residuals and
returns the $\chi^2$ value, the degrees of freedom, and p-values for
each individual covariate along with a global value. Using an
$\alpha = 0.05$ and the global p-value, the determination of whether to
reject or fail to reject $H_0$ can be made. In addition to the hypothesis test, these Shoenfeld residuals can be plotted against time for each covariate to visualize the assumption. When plotting the residuals against time, if the assumption is not broken it would be expected that the points show roughly a straight horizontal line to indicate constant hazard over time [@nahhas2025].

@@

Below is an example plot demonstrating Shoenfeld residuals plotted against time to test the proportional hazards assumption. While there is some curvature in the line, the residuals are roughly horizontal indicating proportional hazard over time for this covariate. In a model with multiple covariates, each one would need to be visualized to ensure it does not break this assumption and can be used in the model.

```{r, echo=FALSE}
example_surv_object <- Surv(time = lung$time, event = lung$status)

example_coxmodel <- coxph(example_surv_object ~ meal.cal + age, data = lung)

example_phtest <- cox.zph(example_coxmodel)

plot(example_phtest[1],
     ylab = "Covariate Schoenfeld Residual",
     xlab = "Time")
```

@@

Below is an example plot showing Martingale residuals plotted against the range of values in the covariate being tested. From the plot, it is clear there is no clear slope in the residuals indicating the linearity assumption is not broken. For a model with multiple covariates, multiple plots will need to be made.

```{r,echo=FALSE}
example_martingale <- resid(example_coxmodel, type = "martingale")

plot(example_martingale, 
     main = "Martingale Residuals: Example Covariate",
     ylab = "Martingale Residuals for Example Covariate",
     xlab = "Value Range for Example Covariate")
abline(h = 0, col = "red")
```

@@

A forest plot will be created using the `plot_model()` function from the sjPlot package. The CPH model can be passed as the sole argument to visualize the hazard ratios. An example forest plot is shown below plotting the hazard ratio for each covariate in the model. The point represents the calculated hazard ratio and the whiskers show the 95% confidence interval.

```{r, echo=FALSE,warning=FALSE}
plot_model(example_coxmodel,
           dot.size = 1,
           line.size = 1,
           colors = "red")+
  theme_bw()+
  labs(title = "Estimated Effects of Covariates on Hazard Ratios")+
  ylab(label = "Covariates")+
  xlab(label = "Estimated Effect on Hazard Ratio")
```

@@

After generating the CPH model, a Kaplan-Meier survival curve can be generated to view survival probabilities against time. These plots are important when modeling survival data because they clearly demonstrate the time where the event being modeled for is expected to occur.
'
The survival curve will be plotted by using the survival package's `survfit()` function. The object created will be plotted using `plot()`

Below is an example survival curve showing survival probabilities over time. The line represents the median survival probability with the red area around it showing the 95% confidence interval for the values. The curve can be interpreted as showing the probability of event would be expected to happen around a time of 300 where the survival probability is below 50%.

```{r, echo=FALSE}
example_curve <- survfit(example_coxmodel)

ggsurvplot(example_curve,
           data = lung,
           title = "Example Survival Curve",
           xlab = "Time",
           ylab = "Survival Probability",
           legend = "none")

@@

As defined by the formula, cumulative hazard is a summation of the hazard rates for an individual from time 0 to time $t$. A higher cumulative hazard indicates a higher risk of event. Cumulative hazard is used as an additional measure for risk, but it is mainly used to calculate survival probability to plot against time on survival curves. [@klein2005]

The cumulative hazard will be calculated in R by grouping the data by the id field and finding the summation of the hazard rates for the individual through the duration of the observation period. A column will be added showing the cumulative hazard for each individual at each point in time.

@@

The resulting survival probability will be between 0 and 1 with a value close to one indicating a high probability of the individual not experiencing the event. Conversely, a value close to 0 indicates a higher probability of the individual experiencing the event. A survival probability of .05 indicates that only 5% of individuals under the same conditions would survive. Survival probability is an important metric when modeling, visualizing, and making predictions with survival data.[@klein2005]

@@

@@

```{r}
survfit <- survfit(cox)

ggsurvplot(survfit, data = test)
```
